### 2020-06-08 doc/response_2_reviews_Ecography2.txt
revisions round 2

Recommendation by the Subject Editor (anonymous):

Both reviewers found the manuscript to be much improved from the previous version, I fully agree with them. While reviewer 1 has only minor comments, reviewer 2 still has some more major concerns that need to be addressed. Thus I invite you to prepare a revision which implements those suggestions.

Reviewer(s)' Comments to Author:
Reviewer: 1

Comments to the Author
The manuscript has improved since its previous round. The work is interesting and carefully executed even though authors should be a bit more critical on the generality of their results. After all this is a very theoretical study where the simulation model is rather simplistic description of the reality.

ADD CAVEATS?

Few specific points:
- space-for-time model. I don't understand the motivation behind this term. Moreover, I don't think this is a commonly used term in ecology. I would encourage revising it.

RESPONSE: We responded to this comment in our first revision, pointing out that "a quick search on Google Scholar shows many papers with "space-for-time" in the title or abstract, including papers cited hundreds of times. Stuart Pickett's 1989 book chapter in the Long-term Studies in Ecology, titled, "Space-for-Time Substitution as an Alternative to Long-Term Studies," has been cited almost 1300 times (Google Scholar

- On line 254, figures and elsewhere. Please provide units (year?) for time

RESPONSE: We added text to this line, and to most figure legends, to clarify that time is measured in years.

Reviewer: 2

Comments to the Author
I reviewed this previously, and asked that more details be included about the methods. Whilst there is some improvement, it is still not up to the standard I would hope for. For example, the text still does not say which years are used in the estimation. It is now clear that only a single simulation is used, which is not really acceptable. We need to be shown (not just told) that the results replicate. being able to replicate the simulations and analysis would also make it easier to run the models under different conditions. This would enable the exploration of a wider range of conditions, which should help to explain the main results.

I would recommend finding a single statistic that can be used to summarise the predictive success, e.g. the probability of the data given the predictions. This can be calculated across some or all of the sites (rather than having to chose a single site to look at). So, for a single simulation, data can be taken at time t, with spatial data across the whole region, and the same amount of temporal data (from either one or several sites). The predictions can be compared to the "real" data at (say) 10 or 1000 years into the future. This basic set-up can be replicated repeatedly under different scenarios, e.g. no climate change, climate change before the sampling, climate change after the sampling, etc.

RESPONSE: We agree, "showing" is better than "telling." We chose not to take the reviewer's suggestion of finding a single measure of model accuracy, because we want readers to focus on the most important general result, which is the shift in model weight from time-series to space-for-time with increasing forecast time scale. We have added a new supplementary figure (SI Fig. 1) showing the temporal trend in the model weight term for 10 independent realizations of the community tunover example and the eco-evolutionary example. We believe this figure shows how consistent the pattern is across simulation runs.

One issue that I think needs to be addressed is the problem of extrapolation outside of the environmental conditions when the sampling is done. In practice, this is a major concern. If, say, methods work well until they extrapolate, then that would be an important message for studies of real organisms.

RESPONSE: We added text to the discussion to flag the important issue of extrapolation. Third to last discussion paragraph now reads as follow, where \comments{} show the new text. 
"The idea of combining forecasts with model weights and allowing the weights to shift across the forecast horizon need not be limited to \comment{extremely simple statistical models like the ones we used in this study}. The same concept could work for any class of models that differ in predictive skill at different forecast horizons. \comment{One possibility might be more sophisticated phenomenological models designed to minimize problems of extrapolating outside the historical range of variation. A set of process-based models focusing on processes operating at different time scales would be another.}  Determining the model weights \emph{a priori} might be easier when models feature explicit processes with characteristic time scales."
We note that the extrapolation issue is also addressed in the Introduction (lines __).