### 2020-07-07 doc/response_2_reviews_Ecography3.txt
final accepted version

03-Jul-2020

Dear Dr. Adler:

Manuscript ID ECOG-05271.R1 entitled "Matching the forecast horizon with the relevant spatial and temporal processes and data sources" which you submitted to Ecography, has been reviewed.  The comments of the reviewer(s) and recommendation by the Subject Editor are included at the bottom of this letter.

The reviewer(s) and the Subject Editor have recommended some minor revisions to your manuscript.  Therefore, I invite you to respond to the reviewer(s)' comments and revise your manuscript.

Recommendation by the Subject Editor (anonymous):
Your manuscript was evaluated by the reviewer who was the more critical reviewer of the original submission (the more positive reviewer was satisfied already with the revisions made earlier).

This was not an easy decision to make: reviewer 2 recommends rejection, because the same fundamental issues remain that were present basically in the original submission. Based on my own reading, I agree with the reviewer that your study has substantial limitations. At the same time, based on my own reading, I consider that the manuscript makes a very important point, and that it is generally very nicely written. Thus, in my view the models presented are simply "proofs of concepts" that help to discuss the (quite obvious but yet maybe not sufficiently acknowledged) fact that time-series models should do better in short-term predictions and space-for-time models should not better in long-term predictions. Given that making the examples more realistic and general would lead to a completely new manuscript, my options were basically to reject the manuscript, or accept it with these limitations. Given the above, I ended up to a decision that was favorable for your side. Thus, I invite you to prepare a minor revision, where you make one more effort to double-check that you don't overstate the generality of the findings, but rather present your models as proofs of concepts that help to illustrate your verbal argument. Related to this, I agree with the reviewer that you may have overstated the value of the simplistic linear weighting approach (even if you discuss its limitations nicely in the main text), as based on your text some readers might think that "just by figuring out the weighting functions one can make accurate predictions for any times-scale". This is of course not expected to be the case for any real case study, even if the optimal weighting would be known. So I would propose that you discuss even more critically the prospects of the weighted-model approach in the revised version.

RESPONSE: We thank the editor for their favorable decision, and for the opportunity to revise. We have addressed the reasonable concern about the potential value of the model weighting approach by adding a new, short paragraph to the Discussion: "On the other hand, there is no guarantee that our proposed model weighting scheme will work when applied in real ecosystems. The most obvious potential problem is that space-for-time approaches may fail to predict long-term dynamics if model assumptions are violated, transient dynamics are strong, or future environmental conditions have no current analog \citep{Worth2014,Veloz2012}. The notion that model weighting can improve forecasts at intermediate and long forecast horizons must be a viewed as a hypothesis to be tested with empirical data."

RESPONSE CONT'D: We also added caveats to the section at the end of the Introduction where we briefly describe our findings (changes tracked in red font).

Reviewer(s)' Comments to Author:
Reviewer: 2

Comments to the Author
This is a resubmission. It still suffers from the same problem as the previous version - the authors are not willing to explore the methods fully. The authors state they find the weighting interesting, but even this is only explored under one set of conditions. How do we know it will act the same if (for example) there is no change in climate?

RESPONSE: If there is no change in climate, there will be no need for model weighting. As our results show, the time series approach will work fine on its own.  Pointing this out seems unnecessary because the premise of the whole exercise was to predict the impact of environmental change.

I don't agree that the shift in weighting is so important, and I can't see that it is general. Only one specific scenario is used, so what will happen under different scenarios? And what if different models were used for the projections, e.g. adding a quadratic term?

RESPONSE: We are sorry the reviewer sees so little value in the analysis. However, it is not strictly accurate to say that we only explored one scenario. We studied two different model systems, and, at the reviewers' request, we also explored a non-stationary scenario. Perhaps the reviewer believes we are proposing a solution to a problem. Our real objective is to highlight a previously under-appreciated problem and propose one potential, and novel, solution. We believe we have added enough caveats and qualifiers to make this clear.

RESPONSE: As for the comment about adding a quadratic term, the spatial models for both the metacommunity example and the eco-evo example do include a quadratic effect of temperature (Eq. 1 in Appendix A, also shown in Figs. 2A and 4B). While the temporal model for the metacommunity example does not contain a quadratic effect of temperature (Appendix A Eq. 2, and Fig. 2B), the temporal model for the eco-evo example does (Appendix A Eq. 3). We added text in Appendix A to clarify this distinction. The reason the temporal model for the metacommunity example does not include a quadratic term is that the data from the baseline period show a linear relationship between temperature and biomass (Fig. 2B). Even if we fit a quadratic term, the coefficient would be close to zero and would not improve long-term forecasts. This illustrates a key point of the paper: different kinds of data represent different processes operating at different time scales.

A minor point: check that the descriptions of the models match the code. I was wondering why there isn't  a quadratic term on K(t) in the temporal model for biomass, but according to the code there isn't one on the spatial model either.

RESPONSE: We address this comment in the paragraph above. We double-checked our computer code and it is correct. The reviewer may have been looking at the spatial model for total biomass in the community turnover example, rather than the biomass of the focal species. That model does not include a quadratic term because the relationship is monotonic increasing. To minimize the chance of this confusion, we moved this block of code to the end of the script to follow the order in which results are presented.